### 2020-11-30 paper/paper.tex
add paper

\documentclass[11pt]{article}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{apacite}
\usepackage{caption}
\usepackage{subcaption}

% \usepackage[
% backend=biber,
% style=alphabetic,
% citestyle=apa
% ]{biblatex}

% \addbibresource{references.bib} % imports bibliography file
\graphicspath{ {./images/}} % Import images

\title{% 
    Validate Structural Analysis \\ 
    \large Testing GPV Method in First-Price Reverse Auctions}

\author{Qingbo Liu\\[0.5cm]{\small Advisor: Timothy P. Hubbard}}

% \doublespacing
\begin{document}
\maketitle

\section{Introduction}
% previous version  
% Auctions are a price-discovering mechanism in environments where there is no 
% well established known price for items. Participants at auctions have heterogenous
% valuations about the items being auctioned. The standard economic assumption is that 
% bidders draw independent or affalidated valuations before the auction starts and 
% submit their bids as a possibly 
% non-linear function of their valuations and preferences. At the end, winners 
% are determined according to certain rules. At the momement the winner is designated, 
% a particular objective of the auction mechanism that is employed is achieved, 
% such as social welfare maximization, revenue maximiaztion, etc.

% However, there are certain cases where the theory is unlikely to match the reality. 
% For example, in first-price auctions, the usual economic model on bidders' 
% behavior draws on an important concept -- equilibrium. The notion of equilibrium 
% states that if every participant in the auction adopts the equilibrium strategy, 
% no one can gain more advantage by switching to a modified strategy and the 
% equilibrium is achieved usually through repeated playing. The theory 
% predicts that in first-price sealed-bid auctions bidders would bid a 
% fractional part of their values, dependent on the number of participants. 
% The intuition behind the theory is that the larger the number of bidders 
% in a given auction, the more fierce the competition is.  
% The theory is elegant, but the complexities inherent in real life are 
% likely to render the model less predictive than it should suggest.

% The fact that theoretical models might not give reliable predictions 
% has implications for works that rely on them. Structural analysis, 
% which uses statistical tools to infer models's parameters from data and 
% gives predictions based on the data, is especially vulnerable to this attack. 
% For instance, the predicted bidding behavior in first-price sealed-bid auctions 
% might not hold in reality, as demonstrated by the lab study Neugebauer \& Selten (2004). 
% Yet Harrison (1989) also raises methodological critique of the lab experiments 
% in the way of comparison. [more details on this]

% But lab experiment is ultimately different from real auctions and it remains an open 
% problem to empirically verify the effectiveness of theoretical predictions, which 
% is the goal of this paper. This paper tries to address the question of

% what structural analysis enables us to do, its drawbacks and the introduction 
% of machine learning practices to verify the if it is correct

Participants at auctions submit bids for items based on their valuations and 
preferences of the items. There are many ways to think about the transformation from 
valuations and preferences into bids, and the usual economic approach is to consider 
the bids as a mathematical function of valuations and preferences, an approach which 
this paper will take. At the end of auctions, the winner(s) 
is designated based on certain auction rules and the auction data becomes available.
The data includes only bids, the final output of bidding function, and the inputs, 
valuations and preferneces, remain unknown. Yet many interesting and useful analysis 
of auctions entails the konwledge about valuations and preferences. For example, 
to evaluate the performance of auctions, it is necessary to have access to bidder's 
valuations to determine if the particular mechanism has achieved its objective,
which leads to structural estimation.

Structural estimation assumes a particular model and uses statistical methods to 
estimate relevant parameters of the model. The calibrated model can then 
give predictions regarding objects that are beging investigated. For example, 
in first-price sealed-bid reverse auctions, 
one of the mechanisms that the government uses for public project procurements, 
the simple assumption is that all participants independently draw costs from an 
identical cost distribution. 
Each bidder submits a sealed bid to the auctioneer and after all bidders have finished 
submission the winner is the one with lowest bid. With this bidding model, 
structural estimation can be used 
to infer the cost distribution based on observed bids \cite{Guerreetal2000}.
Compared to regression analysis, structural estimation imposes a stringer set of 
assumptions on the data. But more than just discovering correlation between variables 
as regression analysis does, structural estimation is able to produce stronger 
counter-factual predictions, such as the cost distribution as mentioned above.

The central assumption underlying structural estimation is that the model used 
is correct in the particular environment. To overcome such a draback, alternative 
studies such as lab experiments are used to confirm the theory's validity. But 
laboratory environments ultimately differ from the real environments and 
the conclusions from laboratory studies may not hold in more realistic environments. 
This goal of this paper is to test model validity in a new different way by 
borrowing the techniques from statistical learning literature. In statistical 
learning, multiple models for predictive purposes are trained on part of the dataset 
and their performance is evaluated on another reserved split of 
the available dataset. The technique is called cross-validation 
and is used when there are candidate models and their performance needs to 
be compared to select the most fit one. In this paper, I will adapt the idea 
of cross-validation to test the predictiveness of economic models. By spliting 
data into training data and test data as in statistical learning, I will 
run structural estimation on the training data, simulate test data using the 
results from strucutral estimation and compare the simulated data to actual 
test data. If the difference is statistically significant, that suggests 
unfitness of the model in the particular environment. 
In this paper, I will use data from government procurement auctions to test 
first-price sealed-bid equilibrium model \cite{RileySamuelson1981} 
for structural estimation
\footnote{The original goal of this paper also includes comparing the cost 
distribution of DBE/MBE/WBE (standing respectly for Disadvantaged/Minority/Woman
Business Enterprise) to the cost distribution of other firms, which will add 
to the existing literature on understanding how DBE/MBE/WBE perform relative 
to other firms. But due to data availability on these types of firms 
and time constraint on the project as this is a 
term paper, this goal is omitted. }.

% TODO: mention FPSB is the environment we are interested in 
% TODO: also mention why we want the cost distribution: MDB/WDB business

The remainders of the paper are organized as follows. Section 2 discusses the 
characteristics of the environemnt and derives the model for structural 
estimation. Section 3 examines the data and presents estimation results. 
Section 4 runs simulations based on estimation. Section 5 concludes with 
a discussion on the results and possible improvements.

\section{Model}

For the model setup, suppose there are $n$ bidders in the auction, with 
$n \geq 2$, and 
there is a single, indivisible object being auctioned.. Each bidder submits 
a sealed bid based on the an observed cost before the auciton. The 
bidder's cost $c$ is drawn i.i.d from a common distribution $F(\cdot)$ which is 
continuous with density $f(\cdot)$ on the support $[\underline{c}, \overline{c}] \subset \mathbb{R}^{+}$. 
All bidders are quasi-linear utility-maximizers, with the utility function 
$u(c, b) = b - c$. Bidders are therefore assumed to be risk-neutral. 
The number of bidders $n$, the cost distribution $F$ and its support are common 
knowledge to all bidders in the auction. Besides, each bidder $i$ can only observe her 
private cost $c_i$ and no one else's.

Different from first-price sealed-bid auctions where 
the highest bidder wins, in first-price sealed-bid reverse 
auctions the winner is picked as the one who submits the 
lowest bid. In the following subsection, I will derive the 
Bayes-Nash Equilibrium in the FPSB reverse auction environment.
After I have characterized the equilibrium, now I will proceed to describe how 
private costs can be inferred from observed bids based on the equilibrium.

\subsection{Bayes-Nash Equilibrium}
For bidder $i$ in the auction, she needs to report a bid $b_i$ based on 
her private cost $c_i$. The bidder achieves maximal utility 
by maximizing the function 
$$ \max_{b_i} (b_i - c_i)P(\text{win}|b_i), $$ 
which is saying by picking an approriate bid $b_i$, the bidder 
has a good balance between the revenue and probability of winning.

Since this is a first-price reverse auction, where the lowest bidder wins, 
the probability of winning is 
\begin{align*}
    P(\text{win}|b_i) &= P(b_1 \geq b_i, ... b_{i-1} \geq b_i, b_{i+1} \geq b_i, ... b_n \geq b_i) \\
        &= P(b_1 \geq b_i) ... P(b_{i-1} \geq b_i) P(b_{i+1} \geq b_i) ... P(b_n \geq b_i) \\
        &= [1-P(b_1 \leq b_i)] ... [1-P(b_{i-1} \leq b_i)][1-P(b_{i+1} \leq b_i)] ... [1-P(b_n \leq b_i)] \\
        &= \prod_{j=1, j \neq i}^{N}[1-P(b_j \leq b_i)]
\end{align*}

Now, suppose that there is a monotone bidding equilibrium function $s(\cdot)$ such that 
all bidders abide by it and makes their bids based on it, i.e. $b_i = s(c_i)$. 
Then
\begin{align*}
    P(\text{win}|b_i) &= \prod_{j=1, j \neq i}^{N}[1-P(s(c_j) \leq b_i)] \\ 
        &= \prod_{j=1, j \neq i}^{N}[1-P(c_j \leq s^{-1}(b_i))] \\ 
        &= \prod_{j=1, j \neq i}^{N}[1-F(s^{-1}(b_i))] \\ 
        &= [1-F(s^{-1}(b_i))]^{n-1}, 
\end{align*}
where the last equality follows from symmetry -- all bidders draw costs (independently) 
from the same distribution.

The objective function therefore becomes 
$$ \max_{b_i} (b_i - c_i)[1-F(s^{-1}(b_i))]^{n-1}$$ 
and setting the derivative of it with respect to $b_i$ equal to 0 yields
\begin{align}
    [1-F(s^{-1}(b_i))]^{n-1} &= (b_i -c_i)(n-1)[1-F(s^{-1}(b_i))]^{n-2}f(s^{-1}(b_i))\frac{d s^{-1}(b_i)}{d b_i} \nonumber \\
    [1-F(c_i)]^{n-1} &= (s(c_i) -c_i)(n-1)[1-F(c_i)]^{n-2}f(c_i)\frac{1}{s'(c_i)} \nonumber \\
    1 &= (s(c_i)-v_i)(n-1)\frac{f(c_i)}{1-F(c_i)}\frac{1}{s'(c_i)}     \label{eqn:BiddingDiffQ}
\end{align}

The differential equation characterizes the strategy in sealed-bid first-price 
reverse auction and the solution of it, $s(\cdot)$, gives the equilibrium 
bidding function as follows \cite{HubbardPaarsch2009}
\begin{equation}
    s(c) = b =  c + \frac{\int_c^{\overline{c}} [1 - F(u)] du}{[1 - F(c)]^{n-1}}
    \label{eqn:BiddingFunction}
\end{equation}

\subsection{Cost Inference}
Our interest, however, lies not in the particular solution 
of this differential equation, but in using it as an intermediate step 
in deriving cost as a function of bids, i.e. $s^{-1}(\cdot)$.

Assume $G(\cdot)$ represents the cumulative distribution of bids, 
along with the continuous density function $g(\cdot)$ and support 
$[\underline{b}, \overline{b}] = [s(\underline{c}), s(\overline{c})]$.
As in \citeauthor{Guerreetal2000} \citeyear{Guerreetal2000}, by a transformation of random variables, I have the identity 
$G(b) = P(\tilde{b} < b) = P(\tilde{c} < s^{-1}(b)) = F(s^{-1}(b)) = F(c)$
and $g(b) = f(c) /s'(c)$. The ratio of $g(b)/(1 - G(b))$ gives 
$f(c_i)/(1-F(c_i))(1/s'(c_i))$. From this, equation (\ref{eqn:BiddingDiffQ}) becomes 
\begin{equation}
    \label{eqn:ReverseFunction}
    c_i = b_i - \frac{1}{n-1}\frac{g(b_i)}{1-G(b_i)}
\end{equation}

In an environment of $n$ bidders, Equation (\ref{eqn:ReverseFunction}) provides 
the method to estimate psuedo-costs from observed bids with kernel-estimated 
bid density and bid distribution. The estimated costs are psuedo because of  
assumptions that bidders are risk-neutral and play Bayes-Nash equilibrium bidding 
strategy.

\subsection{Estimation}
The estimation consists of two steps. In step 1, I recover costs $c$ based on 
observed bids $b$, by first non-parametrically estimating $G(b)$ and $g(b)$ 
and then using equation (\ref{eqn:ReverseFunction}) to 
generate pseudo-costs. In the second step, I can non-parametrically estimate the cost
density $f(c)$ from pseudo-costs.

The technique used to estimate $G(b)$ and $g(b)$, as well as $f(c)$, is called 
kernel density estimation and has the following form
\begin{equation}
    \hat{f}(x) = \frac{1}{nh}\sum_{i=1}^n K(\frac{x-x_i}{h}),
    \label{eqn:KDE_f}
\end{equation}
and 
\begin{equation}
    \hat{F}(x) = \frac{1}{n}\sum_{i=1}^n I(x_i \leq x), 
    \label{eqn:KDE_F}
\end{equation}
where I have a set of observables ${x_1, x_2, ... x_n}$. However, this  
kernel density estimation (\ref{eqn:KDE_f}) 
method suffers from boundary 
problem and instead I will use the boundary-corrected kernel estimation 
method as proposed in \cite{HickmanHubbard2015}

This two-step estimation method has the advantage that in order to estimate cost 
distribution with observed bids, I do not need to assume a prior distribution 
of $f(c)$ in order to estimate it (which is a parametric method that works by 
first assuming the shape of a distribution and then estimating parameters of the 
distribution from data).

\section{Data and Results}
The bidding data used in this paper is downloaded from the Colorado Department of 
Transportation (https://www.codot.gov/business/bidding). I wil take a short detour of the data itself before 
presenting results on cost estimation.

\subsection{Data Description}
\begin{table}[h!]
    \centering
    \begin{tabular}{|l|l|l|l|l|l|}

    \hline
                                       & Mean         & Median       & \begin{tabular}[c]{@{}l@{}}Standard\\ Deviation\end{tabular} & Max           & Min  \\ \hline
        Total Bid                      & 3,172,164.68 & 1,366,571.20 & 5,491,805.02                                                 & 93,398,000.00 & 0.00 \\ \hline
        \begin{tabular}[c]{@{}l@{}}Percent of Enginner's\\ Estimate (Cost)\end{tabular} & 106.54       & 100.02       & 22.63                                                        & 750.00        & 0.00 \\ \hline
        Number of Bidders              & 5.47         & 5            & 2.5                                                          & 16            & 1    \\ \hline
\end{tabular}
\caption{Basic summary statistics of data}
\label{table:1}
\end{table}

There are in total 1182 auctions in the data. For each auction, I have 
the number of participants $n$, bids submitted by participants and engineer's 
estimate. The summary statistics of data is summarized in Table \ref{table:1}. 
Projects vary from each other in their sizes, types, and many other factors. 
Even the same type of projects will have different costs. As Table \ref{table:1}
shows, the variance in Total Bid is very large and not an ideal candidate for 
cost estimates. Instead, I will work with Percent of Engineer's Estimate to 
infer cost distribution. Percent of Engineer's Estimate is Total Bid divided 
by Engineer's Estimate, which is also known as normalized bids. 
Inferred cost distribution constructed this way will be invariant 
across projects.

\begin{figure}[h]
    \includegraphics[width=0.8\textwidth]{BidFrequency.png}
    \centering
    \caption{Frequency of normalized bids (bids divided by engineer's estimate)}
    \label{fig:BidFrequency}
\end{figure}

Figure \ref{fig:BidFrequency} plots the frequency of normalized bids. Most bids 
are centered around $110\%$, which is slightly higher than engineer's estimate 
with respect to which bids are normalized.
However, there are some outliers located \textit{More} bins, 
which would cause numeric instability in estimation. 
For data that is More bins, the kernel-estimated bid distribution 
$G_B(\cdot)$ is close to $1$ and so $1-G_B(\cdot)$ close to 0. As in 
the right-hand side of equation (\ref{eqn:ReverseFunction}), the 
part $\frac{g(b)}{1-G_B{b}}$ will tend to infinity as $1-G_B(b)$ 
approaches zero, resulting in the estimated cost to be negative. 
The problem will persist over the boundary of the data and is 
inherent in the kernel distribution estimation.
The boundary-correction technique introduced in \cite{HickmanHubbard2015}
applies only to the kernel density estimation, not kernel 
distribution estimation, and overcoming 
the boundary problems in kernel distribution estimation is outside the scope 
of this paper. For simplicity, therefore, I will first use equation 
(\ref{eqn:ReverseFunction}) to estimate costs and then trim the negative costs.

\subsection{Results}
With data comprised of auctions that have different number of participants, 
the general estimation procedure remains the same. But the equation 
(\ref{eqn:ReverseFunction}) that assumes a fixed $n$ for number 
of participants is modified to 
accomodate varying $n$ \cite{Matthewetal2018}
\begin{equation}
    c_{in} = b_{in} - \frac{1}{n-1} \cdot \frac{g(b_{in}|N = n)}{1-G_{B}(b_{in}|N=n)},
    \label{eqn:ReverseFunction_N}
\end{equation}
where $c_{in}$ and $b_{in}$ refer to the cost and bid in $i$-th auction 
with $n$ participants, and the density $g(\cdot|N=n)$ and distribution
$G(\cdot|N=n)$ are estimated on auction data with $n$ participants for 
each $n$ in the dataset.

\begin{figure}[h]
    \includegraphics[width=0.75\textwidth]{cost_bid.jpg}
    \centering
    \caption{Estimated cost plotted against bid.}

    \label{fig:cost_bid}
\end{figure}

Figure \ref{fig:cost_bid} plots the estimated cost against bids in the plot. 
The relationship between costs and bids is almost linear, suggesting that for 
each bid, the inferred cost is very close in terms of numeric value. This is so 
mainly due to the bid distribution. As shown in Figure \ref{fig:BidFrequency}, 
most of the bids are spread in the range [70, 200], which means that for each 
bid $b$, the estimated density value at the specific bid value, $f_c(b)$, is 
small: the peak of $f_c(\cdot)$ being $0.0222$. A closer look at the bid 
distributions of auctions with different number of participants, $n$, reveals 
that the majority of auction data with small $n$, where the competition is less 
fierce and bidders are able to have a larger markup over the costs than those 
having more competitors, bids are mostly distributed around $110\%$. The 
auctions that more spread-out data, however, have large $n$.

\begin{figure}[h]
\centering
    \begin{minipage}[h]{0.5\textwidth}
        \centering
        \includegraphics[width=1\linewidth]{costDensity.jpg}
    \end{minipage}%
    \begin{minipage}[h]{0.5\textwidth}
        \centering
        \includegraphics[width=1\linewidth]{costDistribution.jpg}
    \end{minipage}
\caption{Pseudocost density (on the left) and cost distribution (on the right)}
\label{fig:CostEstimation}
\end{figure}

Figure \ref{fig:CostEstimation} shows the kernel-estimated density and 
distribution of psuedo-costs. Because psuedo-costs are close to corresponding 
bids, the density figure on the left of Figure \ref{fig:CostEstimation} is very 
similar to the histogram of bids, Figure \ref{fig:BidFrequency}. As both plots in 
Figure \ref{fig:CostEstimation} show, the inferred costs also center around 
the range [70, 200].

\section{Simulation}
To test if the estimated cost distribution captures the underlying cost 
distribution that drives observed bids, I will proceed to the second step 
of the adapted cross-validation: simulate bids using estimated cost distribution 
$F_c(\cdot)$ and comparing simulated bids to actual bids. For the test data, 
I will use an auction from the dataset which has 16 bidders. The data is not used 
to estimate the cost distribution because the total number of bids with 
$16$ bidders is less than $32$.

The simulation technique is known as inversion method and runs as follows.
I will draw 16 random numbers uniformly 
distributed on the range $(0 ,1)$ and use the inverse distribution function, 
$F_c^{-1}(\cdot)$ to find the corresponding costs. The bids are then computed 
using equation (\ref{eqn:BiddingFunction}) and the estimated cost distribution 
$F_c$. This process is repeated for 1000 times and the averages are taken.

\begin{table}
    \begin{tabular}{llllll}
    \hline
    Normalized Bids     & Mean                    & Median                   & \begin{tabular}[c]{@{}l@{}}Standard\\ Deviation\end{tabular} & Max                           & Min                          \\ \hline
    Test data           & 91.72\% & 73.11\%       & 0.355\%                  & 178.43\% & 58.74\% \\ \hline
    Simulated Bids      & $1.487 \cdot 10^{46}\%$   & $1.758 \cdot 10^{38}\%$    & $5.759 \cdot 10^{46}\%$                               & $2.230 \cdot 10^{47}\%$         & $7.811 \cdot 10^{32}\%$        \\ \hline
    \end{tabular}
    \caption{Summary Statisticals of test data and simulated dataset}
    \label{table:2}
\end{table}

Table \ref{table:2} summarizes the test data and simulated bids, respectively. The 
anamoly of the simulated data comes from the fact the 15th power of $1-F_c(c)$, a
number that is less than 1, is tiny and so $\frac{1}{1-F_c(c)}$ is large,
inflating the reuslt by an astronomical magnitude
\footnote{This paper is still a work in progress and I am investigating the 
reason why the result is so. More updates later}.

\section{Conclusion}
This paper tries to propose a way to validate models used in structural analysis 
by cross-validation. The dataset is split into training dataset and test 
dataset.
First, this paper has conducted structural estimation on a government 
procurement auction dataset about highway constructions and estimated the 
cost distribution on the training data using a two-step nonparametric procedure.
Second, using the 
estimated cost distribution, the paper then constructs simulated bids with the 
inversion method and compare the simulation to test data. 
(Even though there are no meaningful results yet, it is very likely that 
the two will have statistically significant differnece). 
The statistical difference between the simulation and test data suggests that 
the first-price sealed-bid equilibrium model introduced in \citeauthor{RileySamuelson1981} 
\citeyear{RileySamuelson1981} may not be a suitable model to predict bidder's 
behavior in real-world auctions.

However, in the second step of the cross-validation procedure, comparsing bids 
directly to test theory's validity may not be a good idea, as suggested in 
\citeauthor{Harrison1989} \citeyear{Harrison1989}. \citeauthor{Harrison1989}
\citeyear{Harrison1989} instead proposes to "evaluate subject behavior in the 
expected payoff space". This can be incorporated into the project's future work.

The idea of cross-validation has been an old idea in the machine learning 
literature as a way to compare performance across models. I therefore consider 
it a natural idea to introduce it into structural analysis. Structural estimation 
relies on the model's correctness in modeling agent's behavior, but the verification 
of models is left to lab experiments, not empirical data which holds the final 
authority on models. Hopefully, this paper's experiment would shed light on 
possible directions that structural analysis can head to become more robust.

\medskip

\bibliographystyle{apacite}
\bibliography{references}

\end{document}

### 2020-12-01 paper/paper.tex
fix typos

@@
 \author{Qingbo Liu\\[0.5cm]{\small Advisor: Timothy P. Hubbard}}
+\date{Nov 30, 2020}
 
 % \doublespacing
+\onehalfspacing
 \begin{document}
@@
 is designated based on certain auction rules and the auction data becomes available.
-The data includes only bids, the final output of bidding function, and the inputs, 
-valuations and preferneces, remain unknown. Yet many interesting and useful analysis 
-of auctions entails the konwledge about valuations and preferences. For example, 
+The data includes only bids, the final output of the bidding function, and the inputs, 
+valuations and preferences, remain unknown. Yet interesting and useful analysis 
+of auctions entails the knowledge about valuations and preferences. For example, 
 to evaluate the performance of auctions, it is necessary to have access to bidder's 
@@
 estimate relevant parameters of the model. The calibrated model can then 
-give predictions regarding objects that are beging investigated. For example, 
+give predictions regarding objects that are being investigated. For example, 
 in first-price sealed-bid reverse auctions, 
@@
 Each bidder submits a sealed bid to the auctioneer and after all bidders have finished 
-submission the winner is the one with lowest bid. With this bidding model, 
+submission the winner is the one with the lowest bid. With this bidding model, 
 structural estimation can be used 
@@
 Compared to regression analysis, structural estimation imposes a stringer set of 
-assumptions on the data. But more than just discovering correlation between variables 
-as regression analysis does, structural estimation is able to produce stronger 
+assumptions on the data. But more than just discovering the correlation between 
+variables as regression analysis does, structural estimation is able to produce stronger 
 counter-factual predictions, such as the cost distribution as mentioned above.
@@
 The central assumption underlying structural estimation is that the model used 
-is correct in the particular environment. To overcome such a draback, alternative 
+is correct in a particular environment. To overcome such a drawback, alternative 
 studies such as lab experiments are used to confirm the theory's validity. But 
@@
 the conclusions from laboratory studies may not hold in more realistic environments. 
-This goal of this paper is to test model validity in a new different way by 
+The goal of this paper is to test model validity in a new different way by 
 borrowing the techniques from statistical learning literature. In statistical 
@@
 and is used when there are candidate models and their performance needs to 
-be compared to select the most fit one. In this paper, I will adapt the idea 
-of cross-validation to test the predictiveness of economic models. By spliting 
+be compared to select the fittest one. In this paper, I will adapt the idea 
+of cross-validation to test the predictiveness of economic models. By splitting 
 data into training data and test data as in statistical learning, I will 
 run structural estimation on the training data, simulate test data using the 
-results from strucutral estimation and compare the simulated data to actual 
+results from strucutral estimation, and compare the simulated data to actual 
 test data. If the difference is statistically significant, that suggests 
@@
 In this paper, I will use data from government procurement auctions to test 
-first-price sealed-bid equilibrium model \cite{RileySamuelson1981} 
+the first-price sealed-bid equilibrium model \cite{RileySamuelson1981} 
 for structural estimation
 \footnote{The original goal of this paper also includes comparing the cost 
-distribution of DBE/MBE/WBE (standing respectly for Disadvantaged/Minority/Woman
+distribution of DBE/MBE/WBE (standing respectively for Disadvantaged/Minority/Woman
 Business Enterprise) to the cost distribution of other firms, which will add 
@@
 to other firms. But due to data availability on these types of firms 
-and time constraint on the project as this is a 
+and time constraints on the project as this is a 
 term paper, this goal is omitted. }.
 
-% TODO: mention FPSB is the environment we are interested in 
-% TODO: also mention why we want the cost distribution: MDB/WDB business
-
-The remainders of the paper are organized as follows. Section 2 discusses the 
-characteristics of the environemnt and derives the model for structural 
+The remainder of the paper is organized as follows. Section 2 discusses the 
+characteristics of the environment and derives the model for structural 
 estimation. Section 3 examines the data and presents estimation results. 
@@
 $n \geq 2$, and 
-there is a single, indivisible object being auctioned.. Each bidder submits 
-a sealed bid based on the an observed cost before the auciton. The 
+there is a single, indivisible object being auctioned. Each bidder submits 
+a sealed bid based on an observed cost before the auction. The 
 bidder's cost $c$ is drawn i.i.d from a common distribution $F(\cdot)$ which is 
@@
 the highest bidder wins, in first-price sealed-bid reverse 
-auctions the winner is picked as the one who submits the 
+auctions the winner is the one who submits the 
 lowest bid. In the following subsection, I will derive the 
@@
 $$ \max_{b_i} (b_i - c_i)P(\text{win}|b_i), $$ 
-which is saying by picking an approriate bid $b_i$, the bidder 
+which is saying by picking an appropriate bid $b_i$, the bidder 
 has a good balance between the revenue and probability of winning.
@@
 In an environment of $n$ bidders, Equation (\ref{eqn:ReverseFunction}) provides 
-the method to estimate psuedo-costs from observed bids with kernel-estimated 
-bid density and bid distribution. The estimated costs are psuedo because of  
-assumptions that bidders are risk-neutral and play Bayes-Nash equilibrium bidding 
+the method to estimate pseudo-costs from observed bids with kernel-estimated 
+bid density and bid distribution. The estimated costs are pseudo because of  
+assumptions that bidders are risk-neutral and play a Bayes-Nash equilibrium bidding 
 strategy.
@@
 method suffers from boundary 
-problem and instead I will use the boundary-corrected kernel estimation 
+problem and instead, I will use the boundary-corrected kernel estimation 
 method as proposed in \cite{HickmanHubbard2015}
@@
 distribution with observed bids, I do not need to assume a prior distribution 
-of $f(c)$ in order to estimate it (which is a parametric method that works by 
+of $f(c)$ to estimate it (which is a parametric method that works by 
 first assuming the shape of a distribution and then estimating parameters of the 
@@
 The bidding data used in this paper is downloaded from the Colorado Department of 
-Transportation (https://www.codot.gov/business/bidding). I wil take a short detour of the data itself before 
+Transportation (https://www.codot.gov/business/bidding). I will take a short detour of the data itself before 
 presenting results on cost estimation.
@@
     \begin{tabular}{|l|l|l|l|l|l|}
-
     \hline
@@
         Total Bid                      & 3,172,164.68 & 1,366,571.20 & 5,491,805.02                                                 & 93,398,000.00 & 0.00 \\ \hline
-        \begin{tabular}[c]{@{}l@{}}Percent of Enginner's\\ Estimate (Cost)\end{tabular} & 106.54       & 100.02       & 22.63                                                        & 750.00        & 0.00 \\ \hline
+        \begin{tabular}[c]{@{}l@{}}Percent of Engineer's\\ Estimate (Cost)\end{tabular} & 106.54       & 100.02       & 22.63                                                        & 750.00        & 0.00 \\ \hline
         Number of Bidders              & 5.47         & 5            & 2.5                                                          & 16            & 1    \\ \hline
@@
 There are in total 1182 auctions in the data. For each auction, I have 
-the number of participants $n$, bids submitted by participants and engineer's 
-estimate. The summary statistics of data is summarized in Table \ref{table:1}. 
+the number of participants $n$, bids submitted by participants and the engineer's 
+estimate. The summary statistics of the data are summarized in Table \ref{table:1}. 
 Projects vary from each other in their sizes, types, and many other factors. 
@@
 Figure \ref{fig:BidFrequency} plots the frequency of normalized bids. Most bids 
-are centered around $110\%$, which is slightly higher than engineer's estimate 
+are centered around $110\%$, which is slightly higher than the engineer's estimate 
 with respect to which bids are normalized.
-However, there are some outliers located \textit{More} bins, 
+However, there are some outliers located in the \textit{More} bin, 
 which would cause numeric instability in estimation. 
@@
 \subsection{Results}
-With data comprised of auctions that have different number of participants, 
+With data comprised of auctions that have a different number of participants, 
 the general estimation procedure remains the same. But the equation 
-(\ref{eqn:ReverseFunction}) that assumes a fixed $n$ for number 
+(\ref{eqn:ReverseFunction}) that assumes a fixed $n$ for the number 
 of participants is modified to 
-accomodate varying $n$ \cite{Matthewetal2018}
+accommodate varying $n$ \cite{Matthewetal2018}
 \begin{equation}
@@
     \centering
-    \caption{Estimated cost plotted against bid.}
+    \caption{Estimated costs plotted against bids.}
 
@@
 The relationship between costs and bids is almost linear, suggesting that for 
-each bid, the inferred cost is very close in terms of numeric value. This is so 
+each bid, the inferred cost is very close in terms of numeric values. This is so 
 mainly due to the bid distribution. As shown in Figure \ref{fig:BidFrequency}, 
@@
 small: the peak of $f_c(\cdot)$ being $0.0222$. A closer look at the bid 
-distributions of auctions with different number of participants, $n$, reveals 
+distributions of auctions with a different number of participants, $n$, reveals 
 that the majority of auction data with small $n$, where the competition is less 
@@
 having more competitors, bids are mostly distributed around $110\%$. The 
-auctions that more spread-out data, however, have large $n$.
+auctions that more spread-out data, however, have large $n$, in which case 
+the markup over costs is decreased due to the competition.
 
@@
 Figure \ref{fig:CostEstimation} shows the kernel-estimated density and 
-distribution of psuedo-costs. Because psuedo-costs are close to corresponding 
+distribution of pseudo-costs. Because pseudo-costs are close to corresponding 
 bids, the density figure on the left of Figure \ref{fig:CostEstimation} is very 
@@
 
-\section{Simulation}
+\section[title]{Simulation 
+\footnote{This paper is still a work in progress and I am investigating the 
+reason why the result is so. For the latest version on this section, please 
+see \url{https://github.com/QuRyu/CDOT/blob/main/paper/paper.pdf }} } 
 To test if the estimated cost distribution captures the underlying cost 
@@
 
-The simulation technique is known as inversion method and runs as follows.
+The simulation technique is known as the inversion method and runs as follows.
 I will draw 16 random numbers uniformly 
@@
 using equation (\ref{eqn:BiddingFunction}) and the estimated cost distribution 
-$F_c$. This process is repeated for 1000 times and the averages are taken.
+$F_c$. This process is repeated 1000 times and the averages are taken.
 
@@
 Table \ref{table:2} summarizes the test data and simulated bids, respectively. The 
-anamoly of the simulated data comes from the fact the 15th power of $1-F_c(c)$, a
+anomaly of the simulated data comes from the fact the 15th power of $1-F_c(c)$, a
 number that is less than 1, is tiny and so $\frac{1}{1-F_c(c)}$ is large,
-inflating the reuslt by an astronomical magnitude
-\footnote{This paper is still a work in progress and I am investigating the 
-reason why the result is so. More updates later}.
+inflating the result by an astronomical magnitude.
 
@@
 This paper tries to propose a way to validate models used in structural analysis 
-by cross-validation. The dataset is split into training dataset and test 
-dataset.
+by cross-validation. The dataset is split into training data and test 
+data.
 First, this paper has conducted structural estimation on a government 
@@
 (Even though there are no meaningful results yet, it is very likely that 
-the two will have statistically significant differnece). 
+the two will have statistically significant differences). 
 The statistical difference between the simulation and test data suggests that 
@@
 
-However, in the second step of the cross-validation procedure, comparsing bids 
+However, in the second step of the cross-validation procedure, comparing bids 
 directly to test theory's validity may not be a good idea, as suggested in 
@@
 \citeyear{Harrison1989} instead proposes to "evaluate subject behavior in the 
-expected payoff space". This can be incorporated into the project's future work.
+expected payoff space", which can be incorporated into the project's future work.
 
 The idea of cross-validation has been an old idea in the machine learning 
-literature as a way to compare performance across models. I therefore consider 
+literature as a way to compare performance across models. I, therefore, consider 
 it a natural idea to introduce it into structural analysis. Structural estimation 

### 2020-12-06 Extensions/bcgpv_driver.m
fix simulation bug

@@ -24,7 +24,7 @@ set(0,'defaulttextinterpreter','latex','Defaulttextfontsize',fsize);
 % N = 2;
 % bmax = max(b);
 % bmin = min(b);
-% b = sort(b);
+% b = sort(b);f
 
 % use hardle bandwidth transformation constant for non-gaussian kernel
 usehardle = 1;
@@ -136,7 +136,16 @@ fV_bc(vind) = fV_bc_right(vind);
 
 FV = kscdf(v_bc, 'edf', evalpts);
   
-if plotind == 1
+if plotind == 1 
+    % bid histogram 
+    figure 
+    set(gcf,'DefaultLineLineWidth',lwidth)
+    set(gca,'FontSize',fsize)
+    hist(data.PercentOfEstimates, 0:10:200)
+    xlabel('$\mathbf{b}$')
+    ylabel('Frequency')
+    box on 
+    
     % scatter plot of bid function
     figure
     set(gcf,'DefaultLineLineWidth',lwidth)
@@ -174,18 +183,18 @@ if plotind == 1
     box on
 end
 
-
 evalpts = linspace(min(v_bc), max(v_bc), max(v_bc)-min(v_bc));
 FV = kscdf(v_bc, 'edf', evalpts);
 
 if simulind == 1
-    b = [0, 0, 0, 0, 0, 0, 0, 0, 0, 0];
-    N = 2;
+    N = 16;
+    b = zeros(N, 1);
+    c = zeros(N, 1);
     
+    run = 10;
     
-    
-    for i = 1:1000
-        for j = 1:10 
+    for i = 1:run
+        for j = 1:N 
             
             while true 
                 rand_number = rand;
@@ -200,7 +209,7 @@ if simulind == 1
                     end 
                 
                     b(j) = b(j) + cost + ... 
-                        trapz(evalpts(closestIdx:end), 1-FV(closestIdx:end)) / (1-FV(closestIdx))^N;
+                        trapz(evalpts(closestIdx:end), (1-FV(closestIdx:end)).^(N-1)) / (1-FV(closestIdx))^(N-1);
                 else 
                     cost = mean(evalpts(idx));
                     
@@ -209,13 +218,37 @@ if simulind == 1
                     end 
                     
                     b(j) = b(j) + cost + ... 
-                        trapz(evalpts(idx(1):end), 1-FV(idx(1):end)) / (1-prob(i, j))^N;
[…]

### 2020-12-06 paper/paper.tex
finish simulation section

@@
 \author{Qingbo Liu\\[0.5cm]{\small Advisor: Timothy P. Hubbard}}
-\date{Nov 30, 2020}
 
@@
 \begin{equation}
-    s(c) = b =  c + \frac{\int_c^{\overline{c}} [1 - F(u)] du}{[1 - F(c)]^{n-1}}
+    s(c) = b =  c + \frac{\int_c^{\overline{c}} [1 - F(u)]^{n-1} du}{[1 - F(c)]^{n-1}}
     \label{eqn:BiddingFunction}
@@
 To test if the estimated cost distribution captures the underlying cost 
-distribution that drives observed bids, I will proceed to the second step 
-of the adapted cross-validation: simulate bids using estimated cost distribution 
+distribution that induces observed bids, I will proceed to validation stage: 
+simulate bids using estimated cost distribution 
 $F_c(\cdot)$ and comparing simulated bids to actual bids. For the test data, 
@@
 distributed on the range $(0 ,1)$ and use the inverse distribution function, 
-$F_c^{-1}(\cdot)$ to find the corresponding costs. The bids are then computed 
+$F_c^{-1}(\cdot)$, to find the corresponding costs. The bids are then computed 
 using equation (\ref{eqn:BiddingFunction}) and the estimated cost distribution 
@@
 
-\begin{table}
-    \begin{tabular}{llllll}
-    \hline
-    Normalized Bids     & Mean                    & Median                   & \begin{tabular}[c]{@{}l@{}}Standard\\ Deviation\end{tabular} & Max                           & Min                          \\ \hline
-    Test data           & 91.72\% & 73.11\%       & 0.355\%                  & 178.43\% & 58.74\% \\ \hline
-    Simulated Bids      & $1.487 \cdot 10^{46}\%$   & $1.758 \cdot 10^{38}\%$    & $5.759 \cdot 10^{46}\%$                               & $2.230 \cdot 10^{47}\%$         & $7.811 \cdot 10^{32}\%$        \\ \hline
-    \end{tabular}
-    \caption{Summary Statisticals of test data and simulated dataset}
-    \label{table:2}
-\end{table}
+\begin{figure}
+\centering 
+    \begin{minipage}[h]{0.55\textwidth}
+        \centering
+        \includegraphics[width=1\linewidth]{simulated_bids.jpg}
+        \caption{Simulated bids against simulated costs.}
+        \label{fig:simulation}
+    \end{minipage}%
+    \begin{minipage}[h]{0.55\textwidth}
+        \centering
+        \includegraphics[width=1\linewidth]{actualVsimulated.jpg}
+        \caption{Bids from test data (x-axis) plotted 
+        against simulated bids (y-axis).}
+        \label{fig:actualVsimulation}
+    \end{minipage}
+\end{figure}
 
-Table \ref{table:2} summarizes the test data and simulated bids, respectively. The 
-anomaly of the simulated data comes from the fact the 15th power of $1-F_c(c)$, a
-number that is less than 1, is tiny and so $\frac{1}{1-F_c(c)}$ is large,
-inflating the result by an astronomical magnitude.
+Figure \ref{fig:simulation} plos simulated costs against simulated bids that are 
+inferred from simulated costs using equation (\ref{eqn:BiddingFunction}). 
+The simulated costs and bids center around 110\% 
+and the range is small, with the max and min differing by less than 4\%. This 
+is due to the estimated unimodal cost distribution 
+concentrating in the range [80, 200], with the peak around 110\%. 
+With the inversion method repeated for 1000 times, it is natural that 
+the costs and bids fall into the place where most bids and costs are 
+concentrated. With the number of runs curtailed to 100 and 50, as shown in 
+Figure \ref{fig:simulation_run_times}, the bids become more spread out and 
+closer to the bid distribution. This shows that the number of runs is 
+an important factor to determine in roder to have a more accurate simulation 
+but it is outside the scope of this paper. For simplicity, I will keep 
+referencing the data generated with 1000 runs.
+
+\begin{figure}
+\centering
+    \begin{minipage}[h]{0.5\textwidth}
+        \centering 
+        \includegraphics[width=\linewidth]{100run.jpg}
+    \end{minipage}%
+    \begin{minipage}[h]{0.5\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{10run.jpg}
+    \end{minipage}
+\caption{Bid simulation run for 100 times (left) and 10 times(right)}
+\label{fig:simulation_run_times}
+\end{figure}
+
+Figure \ref{fig:actualVsimulation} plots test data against simulated bids. 
+A wide difference between simulation and test data is observed here. 
+This difference, however, is not unexpected for two reasons: 
+(1) the test data, which contains an auciton with 16 bidders, does not conform 
+to the the bid distribution depicted in Figure \ref{fig:BidFrequency}, and 
+(2) I am comparing most likely bids predicted by the estimated cost distribution 
+to a particular auction. Therefore, the difference here not suggest the 
+inaccuracy of the estimated cost distribution or the equilibrium model. 
+Instead, this exercise suggests that we may need to increase the sample size 
+to have a more informative comparison.
 

### 2020-12-06 paper/paper.tex
delete footnote about github

@@
 
-\section[title]{Simulation 
-\footnote{This paper is still a work in progress and I am investigating the 
-reason why the result is so. For the latest version on this section, please 
-see \url{https://github.com/QuRyu/CDOT/blob/main/paper/paper.pdf }} } 
+\section[title]{Simulation} 
 To test if the estimated cost distribution captures the underlying cost 

### 2020-12-06 Extensions/bcgpv_driver copy.m
Revert "fix simulation bug"
This reverts commit bdb182032f294cd6cbb95e451f34dbe3f4ecca55.

@@ -30,10 +30,8 @@ set(0,'defaulttextinterpreter','latex','Defaulttextfontsize',fsize);
 usehardle = 1;
 
 % indicator to turn on plotting
-plotind = 0;
+plotind = 1;
 
-% indicator to turn on simulation
-simulind = 1; 
 
 % choice of kernel---everything else is automated
 kernel = 'triweight';
@@ -69,7 +67,6 @@ data.PercentOfEstimates = data.PercentOfEstimates .* 100;
 
 % parameters 
 v = [];
-bids = [];
 vmin = 999;
 vmax = -999;
 participant_count = unique(data.ParticipantsCount);
@@ -80,10 +77,6 @@ for i = 1:length(participant_count)
     N = participant_count(i); % number of participants
     data_subset = data(data.ParticipantsCount == N, :);
     b = data_subset.PercentOfEstimates;
-    
-    if length(b) < 32
-        continue 
-    end 
     
     bmax = max(b);
     bmin = min(b);
@@ -97,17 +90,16 @@ for i = 1:length(participant_count)
     v_bc = b - (gB_bc)./((1-GB)*(N - 1));
 
     v = vertcat(v, v_bc);
-    bids = vertcat(bids, b);
+    
     
 end 
 
-v_bc = v; 
-
+v_bc = v'; 
 
 % remove inf rows
 inf_idx = find(isinf(v_bc));
 v_bc(inf_idx) = []; 
-bids(inf_idx) = []; 
+data(inf_idx, :) = []; 
 % gB_bc_global(inf_idx) = [];
 
 
@@ -133,7 +125,6 @@ evalpts = linspace(min(v_bc),max(v_bc),neval);
 [fV_bc_right,hv_r] = kspdf_bc(-v_bc,kernel,-evalpts,b0,usehardle);
 vind = find(evalpts >= vmax_bc - hv_r);
 fV_bc(vind) = fV_bc_right(vind);
-
 FV = kscdf(v_bc, 'edf', evalpts);
   
 if plotind == 1
@@ -141,8 +132,8 @@ if plotind == 1
     figure
     set(gcf,'DefaultLineLineWidth',lwidth)
     set(gca,'FontSize',fsize)
-    scatter(v_bc, bids,'.b')
-    xlabel('$\mathbf{\hat{c}}$')
+    scatter(v_bc, data.PercentOfEstimates,'.b')
[…]

### 2020-12-06 paper/paper.tex
add some more

@@
 term paper, this goal is omitted. }.
+
+Similar work has been done in \citeauthor{ChernomazYushioto2019} 
+\citeyear{ChernomazYushioto2019}, where they have compared the accuracy 
+of semiparametric and nonparametric structural estimation methods in an 
+asymetric environment.
 
@@
 
-\begin{figure}[h]
+\begin{figure}
     \includegraphics[width=0.8\textwidth]{BidFrequency.png}
@@
 the boundary problems in kernel distribution estimation is outside the scope 
-of this paper. For simplicity, therefore, I will first use equation 
+of this paper. For simplicity, I will first use equation 
 (\ref{eqn:ReverseFunction}) to estimate costs and then trim the negative costs.
@@
 with $n$ participants, and the density $g(\cdot|N=n)$ and distribution
-$G(\cdot|N=n)$ are estimated on auction data with $n$ participants for 
-each $n$ in the dataset.
+$G(\cdot|N=n)$ are estimated on auction data with $n$ participants.
 
-\begin{figure}[h]
+\begin{figure}
     \includegraphics[width=0.75\textwidth]{cost_bid.jpg}
@@
     \caption{Estimated costs plotted against bids.}
-
     \label{fig:cost_bid}
@@
 Figure \ref{fig:cost_bid} plots the estimated cost against bids in the plot. 
-The relationship between costs and bids is almost linear, suggesting that for 
-each bid, the inferred cost is very close in terms of numeric values. This is so 
+Bidder's markup over costs is infinitesimal. This is so 
 mainly due to the bid distribution. As shown in Figure \ref{fig:BidFrequency}, 
-most of the bids are spread in the range [70, 200], which means that for each 
+most of the bids are scattered in the range [70, 200], which means that for each 
 bid $b$, the estimated density value at the specific bid value, $f_c(b)$, is 
@@
 fierce and bidders are able to have a larger markup over the costs than those 
-having more competitors, bids are mostly distributed around $110\%$. The 
-auctions that more spread-out data, however, have large $n$, in which case 
-the markup over costs is decreased due to the competition.
+having more competitors, has bids compressed around $110\%$. The fact 
+that bids and costs are very close also suggests that the auctions are 
+very efficient in helping the Colorado government save money.
 
-\begin{figure}[h]
+\begin{figure}
 \centering
@@
 Figure \ref{fig:CostEstimation} show, the inferred costs also center around 
-the range [70, 200].
+the range $[70, 200]$. As the density outside the range[70, 200] is very slim, 
+the implication will be that simulation based on this cost distribution will not 
+be very accurate in generating values outside the range, which partially 
+decreases the predictive power of the model.
 
@@
 inaccuracy of the estimated cost distribution or the equilibrium model. 
-Instead, this exercise suggests that we may need to increase the sample size 
-to have a more informative comparison.
+Instead, this exercise suggests that a larger test dataset is needed 
+to deliver more informative results.
 
@@
 data.
-First, this paper has conducted structural estimation on a government 
-procurement auction dataset about highway constructions and estimated the 
+First, this paper conducts structural estimation on a government 
+procurement auction dataset about highway constructions and estimates the 
 cost distribution on the training data using a two-step nonparametric procedure.
@@
 inversion method and compare the simulation to test data. 
-(Even though there are no meaningful results yet, it is very likely that 
-the two will have statistically significant differences). 
-The statistical difference between the simulation and test data suggests that 
-the first-price sealed-bid equilibrium model introduced in \citeauthor{RileySamuelson1981} 
-\citeyear{RileySamuelson1981} may not be a suitable model to predict bidder's 
-behavior in real-world auctions.
+But the wide difference between the simulation and test data does not suggests
+that the first-price sealed-bid equilibrium model introduced in 
+\citeauthor{RileySamuelson1981} \citeyear{RileySamuelson1981}
+is not a suitable model to predict bidder's 
+behavior in real-world auctions. Instead, to draw a firm concludesion requires 
+more data.
 
-However, in the second step of the cross-validation procedure, comparing bids 
+Moreover, in the second step of the cross-validation procedure, comparing bids 
 directly to test theory's validity may not be a good idea, as suggested in 
@@
 \citeyear{Harrison1989} instead proposes to "evaluate subject behavior in the 
-expected payoff space", which can be incorporated into the project's future work.
+expected payoff space", which can be incorporated into future work of the project.
 

### 2020-12-08 paper/paper.tex
fix typos

@@
 run structural estimation on the training data, simulate test data using the 
-results from strucutral estimation, and compare the simulated data to actual 
+results from structural estimation, and compare the simulated data to actual 
 test data. If the difference is statistically significant, that suggests 
@@
 of semiparametric and nonparametric structural estimation methods in an 
-asymetric environment.
+asymmetric environment.
 
@@
 
-This two-step estimation method has the advantage that in order to estimate cost 
+This two-step estimation method has the advantage that to estimate cost 
 distribution with observed bids, I do not need to assume a prior distribution 
@@
 To test if the estimated cost distribution captures the underlying cost 
-distribution that induces observed bids, I will proceed to validation stage: 
+distribution that induces observed bids, I will proceed to the validation stage: 
 simulate bids using estimated cost distribution 
@@
 
-Figure \ref{fig:simulation} plos simulated costs against simulated bids that are 
+Figure \ref{fig:simulation} plots simulated costs against simulated bids that are 
 inferred from simulated costs using equation (\ref{eqn:BiddingFunction}). 
@@
 is due to the estimated unimodal cost distribution 
-concentrating in the range [80, 200], with the peak around 110\%. 
-With the inversion method repeated for 1000 times, it is natural that 
+concentrating in the range [80, 200], with a peak around 110\%. 
+With the inversion method repeated 1000 times, it is natural that 
 the costs and bids fall into the place where most bids and costs are 
@@
 closer to the bid distribution. This shows that the number of runs is 
-an important factor to determine in roder to have a more accurate simulation 
+an important factor to determine to have a more accurate simulation 
 but it is outside the scope of this paper. For simplicity, I will keep 
@@
 This difference, however, is not unexpected for two reasons: 
-(1) the test data, which contains an auciton with 16 bidders, does not conform 
-to the the bid distribution depicted in Figure \ref{fig:BidFrequency}, and 
+(1) the test data, which contains an auction with 16 bidders, does not conform 
+to the bid distribution depicted in Figure \ref{fig:BidFrequency}, and 
 (2) I am comparing most likely bids predicted by the estimated cost distribution 
-to a particular auction. Therefore, the difference here not suggest the 
+to a particular auction. Therefore, the difference here does not suggest the 
 inaccuracy of the estimated cost distribution or the equilibrium model. 
@@
 inversion method and compare the simulation to test data. 
-But the wide difference between the simulation and test data does not suggests
+But the wide difference between the simulation and test data does not suggest
 that the first-price sealed-bid equilibrium model introduced in 
@@
 is not a suitable model to predict bidder's 
-behavior in real-world auctions. Instead, to draw a firm concludesion requires 
+behavior in real-world auctions. Instead, to draw a firm conclusion requires 
 more data.

### 2021-03-11 paper/paper.tex
update paper

@@
 \usepackage{subcaption}
+\usepackage[toc, page]{appendix}
 
@@
 \title{% 
-    Validate Structural Analysis \\ 
-    \large Testing GPV Method in First-Price Reverse Auctions}
+    Validate Nash-Equilibrium Equilibrium Model 
+    in First-Price Reverse Auctions}
 
@@
 
+\begin{abstract}
+    This paper tests the predictive power of Bayes-Nash equilibrium model in 
+    first-price procurement auctions through validation set approach. It 
+    first trains the model on a subset of data using the Guerre-Perrigne-Vuong method 
+    and then generates predictions on validation set using inversion method. 
+    The generated predictions are compared to the actual data from validation set.
+    This paper's result suggests that Bayes-Nash equilibrium model does not 
+    yield accurate predictions but the answer might be more nuanced due to 
+    flaws in the paper's approach. 
+\end{abstract}
+
 \section{Introduction}
@@
 % valuations about the items being auctioned. The standard economic assumption is that 
-% bidders draw independent or affalidated valuations before the auction starts and 
+% bidders draw independent or affiliated valuations before the auction starts and 
 % submit their bids as a possibly 
 % non-linear function of their valuations and preferences. At the end, winners 
-% are determined according to certain rules. At the momement the winner is designated, 
+% are determined according to certain rules. At the moment the winner is designated, 
 % a particular objective of the auction mechanism that is employed is achieved, 
-% such as social welfare maximization, revenue maximiaztion, etc.
+% such as social welfare maximization, revenue maximization, etc.
 
@@
 preferences of the items. There are many ways to think about the transformation from 
-valuations and preferences into bids, and the usual economic approach is to consider 
-the bids as a mathematical function of valuations and preferences, an approach which 
-this paper will take. At the end of auctions, the winner(s) 
+valuations and preferences into bids, and one usual economic approach is to consider 
+the bids as a (hopefully not complex) function of valuations and preferences,
+an approach which this paper will take. At the end of auctions, the winner(s) 
 is designated based on certain auction rules and the auction data becomes available.
-The data includes only bids, the final output of the bidding function, and the inputs, 
-valuations and preferences, remain unknown. Yet interesting and useful analysis 
-of auctions entails the knowledge about valuations and preferences. For example, 
-to evaluate the performance of auctions, it is necessary to have access to bidder's 
-valuations to determine if the particular mechanism has achieved its objective,
-which leads to structural estimation.
+The data includes only bids, the final output of the bidding function, whereas 
+valuations and preferences remain unknown, both of which are inputs to the 
+bidding function. Interesting and useful analysis 
+of auctions would entail additional knowledge about valuations and preferences.
+For example, 
+to compare how well DBE (disadvantaged business enterprise) performs in government 
+procurement relative to other types of business enterprises, besides looking 
+at how much DBE firms bid in auctions, it would also be useful to compare 
+project costs of DBE firms to those of others. As project costs are an unknown,
+to do this essentially means conducting structural estimation on the data.
 
-Structural estimation assumes a particular model and uses statistical methods to 
+What is structural estimation? 
+Structural estimation gives the ability to extract useful information from 
+data, just like regression analysis. The difference lies in that structural 
+analysis assumes a particular model and uses statistical methods to 
 estimate relevant parameters of the model. The calibrated model can then 
-give predictions regarding objects that are being investigated. For example, 
-in first-price sealed-bid reverse auctions, 
-one of the mechanisms that the government uses for public project procurements, 
-the simple assumption is that all participants independently draw costs from an 
-identical cost distribution. 
-Each bidder submits a sealed bid to the auctioneer and after all bidders have finished 
-submission the winner is the one with the lowest bid. With this bidding model, 
-structural estimation can be used 
-to infer the cost distribution based on observed bids \cite{Guerreetal2000}.
-Compared to regression analysis, structural estimation imposes a stringer set of 
-assumptions on the data. But more than just discovering the correlation between 
-variables as regression analysis does, structural estimation is able to produce stronger 
-counter-factual predictions, such as the cost distribution as mentioned above.
+generate predictions on objects of interest. For example, 
+in first-price sealed-bid government procurement auctions, where 
+the bidders who bids the lowest wins,
+one way to do structural analysis is assuming that bidders draw their costs 
+independently and use the same bidding function to make bids. 
+By reverse engineering the bidding function, structural analysis 
+can be used to infer the cost distribution from bids \cite{Guerreetal2000}. 
+On the other hand, compared to regression analysis, 
+structural estimation imposes a stronger set of 
+assumptions on the data. These additional assumptions enable 
+structural estimation to be able to produce 
+counter-factual predictions that are not discoverable under 
+regression analysis.
 
+Yet assumptions structural estimation impose on data also come at a cost.
 The central assumption underlying structural estimation is that the model used 
-is correct in a particular environment. To overcome such a drawback, alternative 
-studies such as lab experiments are used to confirm the theory's validity. But 
-laboratory environments ultimately differ from the real environments and 
-the conclusions from laboratory studies may not hold in more realistic environments. 
+is correct in a particular environment of interest. 
+But it is usually difficult to assess the model's fit in such environments. 
+To overcome such a drawback, alternative 
+studies such as lab experiments are used to confirm the theory's validity. 
+Even though laboratory studies try to simulate real environments as much 
+as possible, laboratory environments ultimately differ from those in real life 
+and so the conclusions drawn from lab experiments do not necessarily hold true.
+
 The goal of this paper is to test model validity in a new different way by 
-borrowing the techniques from statistical learning literature. In statistical 
-learning, multiple models for predictive purposes are trained on part of the dataset 
-and their performance is evaluated on another reserved split of 
-the available dataset. The technique is called cross-validation 
-and is used when there are candidate models and their performance needs to 
-be compared to select the fittest one. In this paper, I will adapt the idea 
-of cross-validation to test the predictiveness of economic models. By splitting 
-data into training data and test data as in statistical learning, I will 
-run structural estimation on the training data, simulate test data using the 
-results from structural estimation, and compare the simulated data to actual 
-test data. If the difference is statistically significant, that suggests 
-unfitness of the model in the particular environment. 
+borrowing techniques from statistical learning literature. In the 
+prediction branch of statistical learning, a common 
+technique to evaluate a model's performance is by training the model 
+on a dedicated training set that is a subset of the available dataset and then 
+tests the model's prediction on the validation set that is independent from the 
+training set.  
+In this paper, I will adapt the idea 
+of validation set approach to investigate how well the equilibrium model 
+performs in first-price procurement auctions. Specifically, I will 
+split bidding data into training data and test data as in statistical learning.
+Then I will 
+run structural estimation on the training data, simulate test data using 
+inversion method, and compare the simulated data to actual 
+test data. If the difference is statistically insignificant, that suggests 
+fitness of the model in the particular environment. 
+Similar work has been done in \citeauthor{ChernomazYushioto2019} 
+\citeyear{ChernomazYushioto2019}, where they have compared the accuracy 
+of semiparametric and nonparametric structural estimation methods in an 
+asymmetric environment.
+
 In this paper, I will use data from government procurement auctions to test 
@@
 term paper, this goal is omitted. }.
-
-Similar work has been done in \citeauthor{ChernomazYushioto2019} 
-\citeyear{ChernomazYushioto2019}, where they have compared the accuracy 
-of semiparametric and nonparametric structural estimation methods in an 
-asymmetric environment.
 
@@
 knowledge to all bidders in the auction. Besides, each bidder $i$ can only observe her 
-private cost $c_i$ and no one else's.
+private cost $c_i$.
 
-Different from first-price sealed-bid auctions where 
-the highest bidder wins, in first-price sealed-bid reverse 
-auctions the winner is the one who submits the 
+The paper focuses on the environment of first-price sealed-bid (FPSB) procurement 
+auctions where the winner is the one who submits the 
 lowest bid. In the following subsection, I will derive the 
 Bayes-Nash Equilibrium in the FPSB reverse auction environment.
-After I have characterized the equilibrium, now I will proceed to describe how 
-private costs can be inferred from observed bids based on the equilibrium.
+After I have characterized the equilibrium, I will proceed to describe how 
+private costs can be inferred from observed bids with the assumption 
+that players are following equilibrium strategy.
 
 \subsection{Bayes-Nash Equilibrium}
-For bidder $i$ in the auction, she needs to report a bid $b_i$ based on 
+For bidder $i$ in the procurement auction, she needs to report a bid $b_i$ based on 
 her private cost $c_i$. The bidder achieves maximal utility 
@@
 which is saying by picking an appropriate bid $b_i$, the bidder 
-has a good balance between the revenue and probability of winning.
+has a good balance between the revenue ($b_i - c_i$) and probability of winning.
 
@@
 $$ \max_{b_i} (b_i - c_i)[1-F(s^{-1}(b_i))]^{n-1}$$ 
-and setting the derivative of it with respect to $b_i$ equal to 0 yields
+and taking the first-order condition with respect to $b_i$ yields
 \begin{align}
@@
 of this differential equation, but in using it as an intermediate step 
-in deriving cost as a function of bids, i.e. $s^{-1}(\cdot)$.
+to derive cost as a function of bids, $s^{-1}(\cdot)$.
 
@@
 The estimation consists of two steps. In step 1, I recover costs $c$ based on 
-observed bids $b$, by first non-parametrically estimating $G(b)$ and $g(b)$ 
-and then using equation (\ref{eqn:ReverseFunction}) to 
-generate pseudo-costs. In the second step, I can non-parametrically estimate the cost
+observed bids $b$, by non-parametrically estimating $G(b)$ and $g(b)$ 
+and using equation (\ref{eqn:ReverseFunction}) to 
+generate pseudo-costs. In step 2, I can then estimate the cost
 density $f(c)$ from pseudo-costs.
@@
 \end{equation}
-where I have a set of observables ${x_1, x_2, ... x_n}$. However, this  
+where ${x_1, x_2, ... x_n}$ are the set of observed values. However, this  
 kernel density estimation (\ref{eqn:KDE_f}) 
 method suffers from boundary 
-problem and instead, I will use the boundary-corrected kernel estimation 
+problem: when the estimator is near the boundary values, the estimate 
+will be downward biased. 
+Instead, I will use the boundary-corrected kernel estimation 
 method as proposed in \cite{HickmanHubbard2015}
 
-This two-step estimation method has the advantage that to estimate cost 
-distribution with observed bids, I do not need to assume a prior distribution 
-of $f(c)$ to estimate it (which is a parametric method that works by 
+This two-step non-parametric estimation method has the advantage that to estimate cost 
+distribution from observed bids, I do not need to assume a prior distribution 
+of $f(c)$, whereas a parametric method works by 
 first assuming the shape of a distribution and then estimating parameters of the 
-distribution from data).
+distribution from data. This helps avoid implications brought by assumed distributions.
 
@@
     \end{minipage}
-\caption{Pseudocost density (on the left) and cost distribution (on the right)}
+\caption{Pseudo-cost density (on the left) and cost distribution (on the right)}
 \label{fig:CostEstimation}
@@
 simulate bids using estimated cost distribution 
-$F_c(\cdot)$ and comparing simulated bids to actual bids. For the test data, 
-I will use an auction from the dataset which has 16 bidders. The data is not used 
-to estimate the cost distribution because the total number of bids with 
-$16$ bidders is less than $32$.
+$F_c(\cdot)$ and comparing simulated bids to actual bids. 
+The validation data comes form an auction with 16 bidders.
 
@@
 using equation (\ref{eqn:BiddingFunction}) and the estimated cost distribution 
-$F_c$. This process is repeated 1000 times and the averages are taken.
+$F_c$. This process is repeated 1000 times and the averages of each bid are taken.
 
@@
 closer to the bid distribution. This shows that the number of runs is 
-an important factor to determine to have a more accurate simulation 
-but it is outside the scope of this paper. For simplicity, I will keep 
+an important factor to determine in order to have a more accurate simulation.
+But determining the best number of runs is outside the scope of this paper.  
+For the rest of this paper, I will keep 
 referencing the data generated with 1000 runs.
@@
 This difference, however, is not unexpected for two reasons: 
-(1) the test data, which contains an auction with 16 bidders, does not conform 
+(1) the validation data, which contains an auction with 16 bidders, does not conform 
 to the bid distribution depicted in Figure \ref{fig:BidFrequency}, and 
@@
 Instead, this exercise suggests that a larger test dataset is needed 
-to deliver more informative results.
+to deliver more informative results and raises the need for a better 
+way to compare simulation data to validation data.
 
 \section{Conclusion}
-This paper tries to propose a way to validate models used in structural analysis 
-by cross-validation. The dataset is split into training data and test 
-data.
-First, this paper conducts structural estimation on a government 
-procurement auction dataset about highway constructions and estimates the 
-cost distribution on the training data using a two-step nonparametric procedure.
+This paper tries to propose a way to test economics models by validation set approach. 
+It splits a government 
+procurement auction dataset about highway constructions into training set and 
+validation set. Then it estimates the 
+cost distribution from training data assuming Bayes-Nash equilibrium model. 
 Second, using the 
-estimated cost distribution, the paper then constructs simulated bids with the 
+estimated cost distribution, it constructs simulated bids with the 
 inversion method and compare the simulation to test data. 
-But the wide difference between the simulation and test data does not suggest
+Even though the wide difference between the simulation and test data suggests
 that the first-price sealed-bid equilibrium model introduced in 
 \citeauthor{RileySamuelson1981} \citeyear{RileySamuelson1981}
-is not a suitable model to predict bidder's 
-behavior in real-world auctions. Instead, to draw a firm conclusion requires 
-more data.
-
-Moreover, in the second step of the cross-validation procedure, comparing bids 
+may be not an accurate model to predict bidder's 
+behavior in such environment, the flaws in this paper's experiment design 
+demand further investigation before any firm conclusion can be drawn. 
+For example, 
+in the second step of the cross-validation procedure, comparing bids 
 directly to test theory's validity may not be a good idea, as suggested in 
@@
 
-The idea of cross-validation has been an old idea in the machine learning 
-literature as a way to compare performance across models. I, therefore, consider 
-it a natural idea to introduce it into structural analysis. Structural estimation 
-relies on the model's correctness in modeling agent's behavior, but the verification 
-of models is left to lab experiments, not empirical data which holds the final 
-authority on models. Hopefully, this paper's experiment would shed light on 
-possible directions that structural analysis can head to become more robust.
+The practice of using validation set to verify model's performance
+has been standard in the statistical learning 
+literature. I consider 
+it a natural extension to introduce it into the economics field as a way to 
+test how models match reality. As Structural estimation 
+relies on the model's correctly modeling agent's behavior,
+I hope the introduction of validation set approach would help with model 
+verification besides lab experiments and would help structural estimation 
+becomes more robust.
 
@@
 
+\pagebreak
+\begin{appendices}
+    Please refer to \url{https://github.com/QuRyu/CDOT/tree/main/Extensions} for code.
+\end{appendices}
+
 \end{document}

### 2021-06-16 paper/paper.tex
rename folder

@@
 \title{% 
-    Validate Nash-Equilibrium Equilibrium Model 
-    in First-Price Reverse Auctions}
+    Validate Bayes-Nash Equilibrium Model 
+    in First-Price Reverse Auctions 
+    \footnote{ All code and data are located at
+    \url{https://github.com/QuRyu/CDOT/} }}
 
 \author{Qingbo Liu\\[0.5cm]{\small Advisor: Timothy P. Hubbard}}
+\date{\vspace{-5ex}}
 
@@
     first-price procurement auctions through validation set approach. It 
-    first trains the model on a subset of data using the Guerre-Perrigne-Vuong method 
+    estimates the equilibrium model on a subset of data using the 
+    Guerre-Perrigne-Vuong method 
     and then generates predictions on validation set using inversion method. 
-    The generated predictions are compared to the actual data from validation set.
-    This paper's result suggests that Bayes-Nash equilibrium model does not 
+    The generated predictions are compared to validation set.
+    The result suggests that Bayes-Nash equilibrium model might not 
     yield accurate predictions but the answer might be more nuanced due to 
@@
 A wide difference between simulation and test data is observed here. 
-This difference, however, is not unexpected for two reasons: 
-(1) the validation data, which contains an auction with 16 bidders, does not conform 
-to the bid distribution depicted in Figure \ref{fig:BidFrequency}, and 
-(2) I am comparing most likely bids predicted by the estimated cost distribution 
-to a particular auction. Therefore, the difference here does not suggest the 
-inaccuracy of the estimated cost distribution or the equilibrium model. 
-Instead, this exercise suggests that a larger test dataset is needed 
-to deliver more informative results and raises the need for a better 
-way to compare simulation data to validation data.
+This difference, however, is not unexpected. 
+The validation data, which contains an auction with 16 bidders, does not conform 
+to the bid distribution depicted in Figure \ref{fig:BidFrequency}. 
+The calibrated model has only learnt information from the training data and 
+so can only make predictions from what it already knows. This might imply that 
+equilibrium model adopted in this paper is not robust to new data. 
+To derive more meaningful comparisons, I need to augment the size of validation set.
+For example, 80\% of available data should be used for training and 20\% of 
+data for validation. 
+However, due to time limit that this term paper imposes, I have not proceeded 
+further.
 
@@
 
-\pagebreak
-\begin{appendices}
-    Please refer to \url{https://github.com/QuRyu/CDOT/tree/main/Extensions} for code.
-\end{appendices}
-
 \end{document}
